---
title: "Select Inshore Recreational Fishing Effort Summary 2025: `r tmplist$county` County"
author:  
- "Prepared for `r tmplist$county` County by Edward Camp (edvcamp@ufl.edu) and Anna Brasawell (a.braswell@ufl.edu), Assistant Professors at the University of Florida`r tmplist$author_sep` `r tmplist$agent``r tmplist$separator` `r tmplist$agent_title` `r tmplist$agent_email`"  
output:
  word_document:
    toc: false  
    pandoc_args:
     '--lua-filter=page-break1.lua'

---

```{r global_controls, echo=FALSE, message=FALSE, warning = FALSE}
##############################################################################
### CONTROL PANEL ###
##############################################################################
  #Spatial 
    #To be Changed
      cntname <- tmplist$county_R_name      
      stname <- tmplist$state
      stname_abrv <- "FL"
      # coastname <- "Atlantic"  #Not applicable for any states other than Florida
      # regname <- "East Central FL"
      # cntregc <- c("Volusia", "`r tmplist$county`")
      coast <-tmplist$coast   #this corresponds to "sub_reg" in the data file, 7= gulf, 6= atl
      
      yrstrt <- tmplist$yrstrt   #1995
 	    yrend <- tmplist$yrend     #2019
 	    years <- seq(yrstrt,yrend,1)
 	    
 	    yrstrt_seas <- tmplist$yrstrt_seas
      yrend_seas <- tmplist$yrend_seas
      years_seas <- seq(yrstrt_seas,yrend_seas,1)
      species <- c("Black drum, Sheepshead, Red drum, Spotted seatrout, Lefteye flounder family, Drum family, Sand seatrout, Sand drum,
      							Seatrout genus, Trout family, Gulf flounder, and Southern flounder")  
```

### How to use this summary
This documents summarizes the estimated marine recreational fishing effort (number of fishing trips) in `r tmplist$county` County that targeted the `r tmplist$fishery_name`. The  species considered as the `r tmplist$spec_name` for this summary were `r species`. This information helps describe the magnitude of this specific fishery and its changes over time and season. 

### What information is included and how was it generated
All of the information here comes from the NOAA Fisheries Marine Recreational Information Program (MRIP). This summary describes fishing effort annually (trends over time, `r tmplist$yrstrt`-`r tmplist$yrend`) and seasonally (average trips per month from `r tmplist$yrstrt_seas`-`r tmplist$yrend_seas`). For annual trends, the information is described by fishing **mode**--the type of fishing trip. There are three modes provided in MRIP: shore-based, for-hire (guided trips, charters, party boats, etc.), and private or rental vessel. The seasonal trends are shown for all modes combined, but greater detail is available in the full report. Data are shown through `r tmplist$yrend`. 

### Seasonal pattens in fishing effort
The seasonal patterns of all trips (regardless of mode) is shown in Figure 1. These proportions are the average of the years `r tmplist$yrstrt_seas`-`r tmplist$yrend_seas`. 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning = FALSE)
```


```{r data, echo=FALSE, message=FALSE, warning = FALSE}
library(data.table)
library(Hmisc)
library(dplyr)
library(survey) 

##############################################################################
### DATA ###
##############################################################################
	### Read in saved data
		#MRIP
			setwd("U:/UF/Extension specific/CFEE/Effort trends/code/Markdown/2025")
	  	load("fl_2000_2024.RData")
	  	dat_a <- fl_2000_2024

			load("FL_1995_1999.RData")
			dat_b <- dt_fl

			dat <- rbind(dat_a, dat_b)


		#US info
		#	setwd("U:/UF/Extension specific/CFEE/Effort trends/code/Markdown/2025") 
  		fips <- read.csv("US_FIPS_Codes.csv", header=T)
  		states <- unique(fips[,c("state", "state_code")])


```


```{r derived_controls, echo=FALSE, message=FALSE, warning = FALSE}
##############################################################################
### DERIVED CONTROL PANEL ###
##############################################################################
    #Derived
      area_near_ocean <- ifelse(coast==7, 3, 1)   #So, FL gulf near ocean is "3", and FL Atl near ocean is "1"
      area_far_ocean <- ifelse(coast==7, 4, 2)    #Also FL gulf far ocean is "4", and FL Atl far ocean is "2"
      counties <- unique(fips[which(fips$state==stname),c("county", "county_code")])
      stcode <- fips[which(fips$state==stname), c("state_code")] [1]
      cntcode <- fips[which(fips$state==stname & fips$county==cntname), c("county_code")] 
      #cntregcode <- as.character(c(fips[which(fips$state==stname & fips$county %in% cntregc), c("county_code")]))
      stcntycode <- paste(stcode, "_", cntcode, sep="")
      ocndistplot <- ifelse(coast=="6", "3mi", "10mi")

```


```{r data_operations, echo=FALSE, message=FALSE, warning = FALSE}
##############################################################################
### DATA OPERATIONS / MANIPULATIONS ###
##############################################################################
	#Filtering
		daty <- dat[year %in% years,]  #Can eventually add other filters in here, like could filter the data by mode, county, etc. 

	#Converting back to data.frame
		dat1 <- as.data.frame(daty)  #Data tables are great, but will take time to learn how to integrate with code, so turning back to data.frame

		#additional data manipulation aspects
      dat1$monthc <- as.character(dat1$month)
      dat1$month_f <- as.factor(ifelse(nchar(dat1$monthc)>1,  dat1$monthc, paste("0", dat1$monthc, sep="")))
      dat1$cntyf1 <- as.factor(ifelse(nchar(dat1$cnty)==1, paste("00", dat1$cnty, sep=""),
      					ifelse(nchar(dat1$cnty)==2, paste("0", dat1$cnty, sep=""), dat1$cnty)))
      dat1$cntyf1[which(is.na(dat1$cntyf1))] <- "099"

  		dat1$mode_f <- as.factor(as.character(dat1$mode_f))
  		dat1$mode_fx <- as.factor(as.character(dat1$mode_fx))
  		dat1$area_x <- as.factor(as.character(dat1$area_x))
  		dat1$cntyf <- as.factor(as.character(dat1$cnty)) 
  		dat1$st_cnty <- as.factor(as.character(paste(dat1$st, dat1$cntyf, sep="_")))
  		dat1$st_cnty_res <- as.factor(as.character(paste(dat1$st_res, dat1$cnty_res, sep="_")))
  		dat1$prim1_common <- capitalize(tolower(dat1$prim1_common))
  			levels(dat1$prim1_common) <- c(levels(dat1$prim1_common), "No Target")
  			dat1$prim1_common[which(dat1$prim1_common =="")] <- "No Target"
  		dat1$prim2_common <- capitalize(tolower(dat1$prim2_common))
  			levels(dat1$prim2_common) <- c(levels(dat1$prim2_common), "No Target")
  			dat1$prim2_common[which(dat1$prim2_common =="")] <- "No Target"
  			
  		dat1sp <- filter(dat1, dat1$prim1_common %in% species | dat1$prim2_common %in% species)
```


```{r functions, echo=FALSE, message=FALSE, warning = FALSE}
##############################################################################
### DATA ANALYSES AND MODELS ###
##############################################################################

#main workhorse function, generally used for county by year
	estfun <- function(df)  
		{

			#creating dom_id for the different groups you want estimates for
			df$dom_id <-c(paste(df$year, df$st, df$sub_reg, sep=""))  #here is the domain id only down to year, but subset for the county of concern
				#if you wanted to go to say, wave, you'd need to add dat1$wave in there too. Example from MRIP code below
					#dataset$dom_id<-c(paste(dataset$year,dataset$wave,dataset$st,dataset$sub_reg, dataset$mode_fx,dataset$area_x,sep=""))
			df$dtrip <- 1 		#Ok, I don' tknow what this is, need to sort that out 

			#survey design function, from MRIP
  			dfpc<-svydesign(ids=~psu_id,strata=~strat_id, weights=~wp_int,nest=TRUE,data=df)  #So...assuming this is the survey design info?
   				options(survey.lonely.psu = "certainty")

   			#survey estimates function, from MRIP
 	 		results<-svyby(~dtrip, ~dom_id, dfpc, svytotal, vartype=c("se","cv"), keep.names=FALSE) #and this shoudl be the engine for estimation?
  				names(results)<-c("Domain","Trips","SE","PSE")

  			#Creating additional variables
  			results$year <- as.numeric(substr(results$Domain,1,4))
  			results$state <- substr(results$Domain,5,6)
  			#results$mode_fx <- substr(results$Domain,8,8)
  			results$county_code <- df$cnty[1]  #returns county
  			results$sub_reg <- df$sub_reg[1]  #returns county
  		results
  		}


#SEASONAL estimating the effort at each month NOT YEAR. Designed for county
    estfun_month <- function(df)
    {

      #creating dom_id for the different groups you want estimates for
      #df$dom_id <-c(paste(df$year, df$st, df$mode_fx, df$month_f, sep=""))  #here is the domain id only down to year, but subset for the county of concern
      df$dom_id <-c(paste(df$cntyf1, df$month_f, sep="")) 
        #if you wanted to go to say, wave, you'd need to add dat1$wave in there too. Example from MRIP code below
          #dataset$dom_id<-c(paste(dataset$year,dataset$wave,dataset$st,dataset$sub_reg, dataset$mode_fx,dataset$area_x,sep=""))
      df$dtrip <- 1     #Ok, I don' tknow what this is, need to sort that out 

      #survey design function, from MRIP
        dfpc<-svydesign(ids=~psu_id,strata=~strat_id, weights=~wp_int,nest=TRUE,data=df)  #So...assuming this is the survey design info?
          options(survey.lonely.psu = "certainty")

        #survey estimates function, from MRIP
      results<-svyby(~dtrip, ~dom_id, dfpc, svytotal, vartype=c("se","cv"), keep.names=FALSE) #and this shoudl be the engine for estimation?
          names(results)<-c("Domain","Trips","SE","PSE")

        #Creating additional variables
        nchars = nchar(results$Domain)
        results$county_code <- as.numeric(substr(results$Domain,1,3))
        #results$year <- as.numeric(substr(results$Domain,1,4))
        #results$state <- substr(results$Domain,5,6)
        #results$mode_fx <- substr(results$Domain,4,4)
        results$month_f <- substr(results$Domain,4,5)
        
        # results$sub_reg <- df$sub_reg[1]  #returns county
      #results
      #patch to deal with some months not being sampled
month <- as.factor(c("01", "02", "03", "04", "05", "06", "07", "08", "09", "10", "11", "12")) 
monthdf <- data.frame(month)
test <- merge(x=results, y=monthdf, by.x="month_f", by.y="month", all.x=TRUE, all.y=TRUE)    #merge them based on year-month
test[,c('Trips', 'SE', 'PSE')] [is.na(test[,c('Trips', 'SE', 'PSE')])] <- 0 #just change the NA's to zeros, which is correct
      results <- test
      results
      }


#SEASONAL estimating the effort at each month NOT YEAR. Designed for county
  	estfun_mode_month <- function(df)
		{

			#creating dom_id for the different groups you want estimates for
			#df$dom_id <-c(paste(df$year, df$st, df$mode_fx, df$month_f, sep=""))  #here is the domain id only down to year, but subset for the county of concern
			df$dom_id <-c(paste(df$cntyf1, df$mode_fx, df$month_f, sep="")) 
				#if you wanted to go to say, wave, you'd need to add dat1$wave in there too. Example from MRIP code below
					#dataset$dom_id<-c(paste(dataset$year,dataset$wave,dataset$st,dataset$sub_reg, dataset$mode_fx,dataset$area_x,sep=""))
			df$dtrip <- 1 		#Ok, I don' tknow what this is, need to sort that out 

			#survey design function, from MRIP
  			dfpc<-svydesign(ids=~psu_id,strata=~strat_id, weights=~wp_int,nest=TRUE,data=df)  #So...assuming this is the survey design info?
   				options(survey.lonely.psu = "certainty")

   			#survey estimates function, from MRIP
 	 		results<-svyby(~dtrip, ~dom_id, dfpc, svytotal, vartype=c("se","cv"), keep.names=FALSE) #and this shoudl be the engine for estimation?
  				names(results)<-c("Domain","Trips","SE","PSE")

  			#Creating additional variables
  			nchars = nchar(results$Domain)
  			results$county_code <- as.numeric(substr(results$Domain,1,3))
  			#results$year <- as.numeric(substr(results$Domain,1,4))
  			#results$state <- substr(results$Domain,5,6)
  			results$mode_fx <- substr(results$Domain,4,4)
  			results$month_f <- substr(results$Domain,5,6)
  			
  			# results$sub_reg <- df$sub_reg[1]  #returns county
  		      #results
        #patch for counties where a mode or area isn't sampled in a given month in the data set
        month <- as.factor(c("01", "02", "03", "04", "05", "06", "07", "08", "09", "10", "11", "12")) 
        monthdf <- data.frame(month)
        test <- merge(x=results, y=monthdf, by.x="month_f", by.y="month", all.x=TRUE, all.y=TRUE)    #merge them based on year-month
        test[,c('Trips', 'SE', 'PSE')] [is.na(test[,c('Trips', 'SE', 'PSE')])] <- 0 #just change the NA's to zeros, which is correct
      results <- test
      results
  		}


#SEASONAL estimating the effort at each month NOT YEAR. Designed for county
    estfun_area_month <- function(df)
    {

      #creating dom_id for the different groups you want estimates for
      #df$dom_id <-c(paste(df$year, df$st, df$mode_fx, df$month_f, sep=""))  #here is the domain id only down to year, but subset for the county of concern
      df$dom_id <-c(paste(df$cntyf1, df$area_x, df$month_f, sep="")) 
        #if you wanted to go to say, wave, you'd need to add dat1$wave in there too. Example from MRIP code below
          #dataset$dom_id<-c(paste(dataset$year,dataset$wave,dataset$st,dataset$sub_reg, dataset$mode_fx,dataset$area_x,sep=""))
      df$dtrip <- 1     #Ok, I don' tknow what this is, need to sort that out 

      #survey design function, from MRIP
        dfpc<-svydesign(ids=~psu_id,strata=~strat_id, weights=~wp_int,nest=TRUE,data=df)  #So...assuming this is the survey design info?
          options(survey.lonely.psu = "certainty")

        #survey estimates function, from MRIP
      results<-svyby(~dtrip, ~dom_id, dfpc, svytotal, vartype=c("se","cv"), keep.names=FALSE) #and this shoudl be the engine for estimation?
          names(results)<-c("Domain","Trips","SE","PSE")

        #Creating additional variables
        nchars = nchar(results$Domain)
        results$county_code <- as.numeric(substr(results$Domain,1,3))
        #results$year <- as.numeric(substr(results$Domain,1,4))
        #results$state <- substr(results$Domain,5,6)
        results$area_x <- substr(results$Domain,4,4)
        results$month_f <- substr(results$Domain,5,6)
        
        # results$sub_reg <- df$sub_reg[1]  #returns county
  		      #results
        #patch for counties where a mode or area isn't sampled in a given month in the data set
        month <- as.factor(c("01", "02", "03", "04", "05", "06", "07", "08", "09", "10", "11", "12")) 
        monthdf <- data.frame(month)
        test <- merge(x=results, y=monthdf, by.x="month_f", by.y="month", all.x=TRUE, all.y=TRUE)    #merge them based on year-month
        test[,c('Trips', 'SE', 'PSE')] [is.na(test[,c('Trips', 'SE', 'PSE')])] <- 0 #just change the NA's to zeros, which is correct
      results <- test
      results
    }
```    


```{r output, echo=FALSE, message=FALSE, warning = FALSE}    
######################################################
#### OUTPUT
######################################################

	### Filters analyses 
		#County by YEAR
			df_filt_cnt_all <- filter(dat1sp, dat1sp$cntyf %in% c(cntcode))  #Filtering for the county in the control panel
			df_filt_cnt_Mshore <- filter(dat1sp, dat1sp$cntyf %in% c(cntcode), dat1sp$mode_fx==3)  #Filtering for the county in the control panel
			df_filt_cnt_Mhire <- filter(dat1sp, dat1sp$cntyf %in% c(cntcode), dat1sp$mode_fx==5)  #Filtering for the county in the control panel
			df_filt_cnt_Mpriv <- filter(dat1sp, dat1sp$cntyf %in% c(cntcode), dat1sp$mode_fx==7)  #Filtering for the county in the control panel

			df_filt_cnt_AnearO <- filter(dat1sp, dat1sp$cntyf %in% c(cntcode), dat1sp$area_x==area_near_ocean)  #Filtering for the county in the control panel
			df_filt_cnt_AfarO <- filter(dat1sp, dat1sp$cntyf %in% c(cntcode), dat1sp$area_x==area_far_ocean)  #Filtering for the county in the control panel
			df_filt_cnt_Ains <- filter(dat1sp, dat1sp$cntyf %in% c(cntcode), dat1sp$area_x==5)  #Filtering for the county in the control panel

      #County by MONTH
      df_filt_cnt_all_seas <- filter(dat1sp, dat1sp$year %in% c(years_seas), dat1sp$cntyf %in% c(cntcode))  #Filtering for the county in the control panel
      df_filt_cnt_Mshore_seas <- filter(dat1sp, dat1sp$year %in% c(years_seas), dat1sp$cntyf %in% c(cntcode), dat1sp$mode_fx==3)  
      df_filt_cnt_Mhire_seas <- filter(dat1sp, dat1sp$year %in% c(years_seas), dat1sp$cntyf %in% c(cntcode), dat1sp$mode_fx==5)  
      df_filt_cnt_Mpriv_seas <- filter(dat1sp, dat1sp$year %in% c(years_seas), dat1sp$cntyf %in% c(cntcode), dat1sp$mode_fx==7)  

      df_filt_cnt_AnearO_seas <- filter(dat1sp, dat1sp$year %in% c(years_seas), dat1sp$cntyf %in% c(cntcode), dat1sp$area_x==area_near_ocean)  
      df_filt_cnt_AfarO_seas <- filter(dat1sp, dat1sp$year %in% c(years_seas), dat1sp$cntyf %in% c(cntcode), dat1sp$area_x==area_far_ocean)  
      df_filt_cnt_Ains_seas <- filter(dat1sp, dat1sp$year %in% c(years_seas), dat1sp$cntyf %in% c(cntcode), dat1sp$area_x==5)  
	

	### COUNTY specific results
		### By YEAR getting results of all, specific modes and areas
			results_all_cnt <- estfun(df_filt_cnt_all)
			results_Mshore_cnt <- estfun(df_filt_cnt_Mshore)
			results_Mhire_cnt <- estfun(df_filt_cnt_Mhire)
			results_Mpriv_cnt <- estfun(df_filt_cnt_Mpriv)
			results_AnearO_cnt <- estfun(df_filt_cnt_AnearO)
			results_AfarO_cnt <- estfun(df_filt_cnt_AfarO)
			results_Ains_cnt <- estfun(df_filt_cnt_Ains)

    ### by MONTH
      results_cnt_seas_all <- estfun_month(df_filt_cnt_all_seas)

      results_Mshore_cnt_seas <- tryCatch({estfun_mode_month(df_filt_cnt_Mshore_seas)}, error=function(e){})
      results_Mhire_cnt_seas <- tryCatch({estfun_mode_month(df_filt_cnt_Mhire_seas)}, error=function(e){})
      results_Mpriv_cnt_seas <- tryCatch({estfun_mode_month(df_filt_cnt_Mpriv_seas)}, error=function(e){})
      results_AnearO_cnt_seas <- tryCatch({estfun_area_month(df_filt_cnt_AnearO_seas)}, error=function(e){})
      results_AfarO_cnt_seas <- tryCatch({estfun_area_month(df_filt_cnt_AfarO_seas)}, error=function(e){})
      results_Ains_cnt_seas <- tryCatch({estfun_area_month(df_filt_cnt_Ains_seas)}, error=function(e){})

      
### Plotting functions
  #Single county plot of all effort (pooled across area and mode)
		plotfun_c <- function (results, main){
			results$lcl <- results$Trips-1.96*results$SE
	  	results$ucl <- results$Trips+1.96*results$SE

			x = results$year
			y = results$Trips 
			ly = results$lcl
			uy = results$ucl 

			ymax <- 1.25*max(uy)
		  ymin <- .90*min(ly)

		#Transparency function
			col2rgbA<-function(color,transparency) {
			  rgb(t(col2rgb(color))/255,alpha=transparency)
			}

			matplot(x=x,y=cbind(y, ly, uy),type="n", xaxt="n", yaxt="n", 
				xlab="Year",ylab="",col="black", ylim=c(ymin,ymax))
			axis(side=1, at=seq(min(x), max(x), by=1))
			axis(side=2, at=pretty(seq(ymin, ymax, by=(ymax-ymin)/6)), labels=formatC(pretty(seq(ymin, ymax, by=(ymax-ymin)/6)), big.mark=",", format="d"), las=1)
			mtext("Trips", side=2, line=5, adj=.5, outer=FALSE)
			polygon(x=c(x,rev(x)),y=c(ly, rev(uy)), col=col2rgbA("lightblue", .5), border = "lightblue")
			points(x, y, type="l", lwd=2, col="dodgerblue3")
			legend("topright", legend=c("Annual mean", "95% Conf. int."), col=c("dodgerblue3", col2rgbA("lightblue", .5)), lwd=c(2, 8), lty=c(1,1), bty="n")
			#text(median(x), 0.97*ymax, paste(cntname, " county", main, sep=""), cex=2) 
			mtext(expression(italic("Source: NMFS MRIP Database")), side=1, line=3, adj=1, outer=FALSE)
			mtext(paste(cntname, " County", main, sep=""), side=3, cex=2, line=0, adj=.5, outer=FALSE)
			}


	#3-panel plot of mode or area-specific effort
					plotfun3_c <- function (results, main){
					results$lcl <- results$Trips-1.96*results$SE
			  		results$ucl <- results$Trips+1.96*results$SE

			  			x = results$year
              x_main = seq(from=min(x), to=max(x), by=1)
			  			y = results$Trips 
			  			ly = results$lcl
			  			uy = results$ucl 

			  			ymax <- 1.25*max(uy)
						  ymin <- .90*min(ly)

						### Transparency function
							col2rgbA<-function(color,transparency) {
							  rgb(t(col2rgb(color))/255,alpha=transparency)
							}

							matplot(x=x,y=cbind(y, ly, uy),type="n", xaxt="n", yaxt="n", 
								xlab="Year",ylab="",col="black", ylim=c(ymin,ymax), xlim=c(min(years), max(years)))
              axis(side=1, at=years)
							axis(side=2, at=pretty(seq(ymin, ymax, by=(ymax-ymin)/6)), labels=formatC(pretty(seq(ymin, ymax, by=(ymax-ymin)/6)), big.mark=",", format="d"), las=1)
							mtext("Trips", side=2, line=5, adj=.5, outer=FALSE)
							polygon(x=c(x,rev(x)),y=c(ly, rev(uy)), col=col2rgbA("lightblue", .5), border = "lightblue")
							points(x, y, type="l", lwd=2, col="dodgerblue3")
							legend("topright", legend=c("Annual mean", "95% Conf. int."), col=c("dodgerblue3", col2rgbA("lightblue", .5)), lwd=c(2, 8), lty=c(1,1), bty="n")
							text(median(x_main), 0.94*ymax, paste(main, sep=""), cex=2) #this works but puts a stupid space in where you don't need it. 
							#mtext(expression(italic("Source: NMFS MRIP Database")), side=1, line=3, adj=1, outer=FALSE)
			}




  #Single SEASONAL county plot of all effort (pooled across area and mode)
    plotfun_c_seas <- function (results, main){
      results_prop <-results$Trips/sum(results$Trips)
      bpnames=c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
      cols= colorRampPalette(c("dodgerblue", "pink"))(12)       #create a vector of colors equal to the number of bins
      cols_ordered = cols[rank(results_prop)]
      color="dodgerblue"

      barplot(results_prop, names.arg=bpnames, col=color, ylim=c(0,1.2*max(results_prop)), ylab="Percent of effort", main="", 
      cex.lab=0.7, cex.axis=0.7, cex.names=0.7)           
          mtext(expression(italic("Source: NMFS MRIP Database")), side=1, cex=0.6, line=2, adj=1, outer=FALSE)
          mtext(paste(cntname, " County", main, sep=""), side=3, cex=0.75, line=0, adj=.5, outer=FALSE)
      }

    plotfun_c_seas3 <- function (results, main){
      results_prop <-results$Trips/sum(results$Trips)
      bpnames=c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
      cols= colorRampPalette(c("dodgerblue", "pink"))(12)       #create a vector of colors equal to the number of bins
      cols_ordered = cols[rank(results_prop)]
      color="dodgerblue"

      barplot(results_prop, names.arg=bpnames, col=color, ylim=c(0,1.2*max(results_prop)), ylab="Percent of effort", main="", 
      cex.lab=1.5, cex.axis=1.25, cex.names=1.5)           
          # mtext(expression(italic("Source: NMFS MRIP Database")), side=1, line=3, adj=1, outer=FALSE)
          mtext(main, side=3, cex=1.25, line=-1, adj=.5, outer=FALSE)
      }

          plotfun_c_seas3_nodata <- function (){
          results_prop <-rep(0,12)
          bpnames=c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
          cols= colorRampPalette(c("dodgerblue", "pink"))(12)       #create a vector of colors equal to the number of bins
          cols_ordered = cols[rank(results_prop)] 
          color="dodgerblue"

          barplot(results_prop, names.arg=bpnames, col=color, ylim=c(0,1.2), ylab="Percent of effort", main="", 
          cex.lab=1.5, cex.axis=1.25, cex.names=1.5)           
              # mtext(expression(italic("Source: NMFS MRIP Database")), side=1, line=3, adj=1, outer=FALSE)
              mtext("No Data Available", side=3, cex=1.25, line=-5, adj=.5, outer=FALSE)
      }
    
```

  
```{r fig1, fig.width=6, fig.height=2, echo=FALSE, message=FALSE, warning = FALSE}
    options(scipen=999)
    par(mfrow=c(1,1), mgp = c(1.5, .25, 0), mar=c(3,4,1,2), oma=c(0,2,0,1))  
    plotfun_c_seas(results=results_cnt_seas_all, main=paste(", all modes and areas, ", yrstrt_seas, "-", yrend_seas, sep=""))
```

**Figure 1. `r tmplist$county` County seasonal patterns of `r tmplist$fishery_name` effort, all modes.** The percent of r tmplist$fishery_name` effort from years `r tmplist$yrstrt_seas`-`r tmplist$yrend_seas` occurring in each month.  

### Annual effort trends by mode
Effort by mode may be important for deciding about things like advertising or allocating funds for access. The trends by mode are shown in Figure 2. The dark blue line is the most likely estimate and the light blue area represents the uncertainty.

```{r fig2, fig.width=9, fig.height=6.5, echo=FALSE, message=FALSE, warning = FALSE}
		par(mfrow=c(3,1), mgp = c(2.5, .5, 0), mar=c(4,4,2,2), oma=c(.5,3,.5,1))  
		options(scipen=999)
		plotfun3_c(results=results_Mshore_cnt, main="Shore-based, all areas")
			mtext(bquote(italic(.(cntname)~"County")), side=3, line=0, adj=0, cex=1.5, outer=FALSE)
		plotfun3_c(results=results_Mhire_cnt, main="For-hire, all areas")
		plotfun3_c(results=results_Mpriv_cnt, main="Private/rental, all areas")
			mtext(expression(italic("Source: NMFS MRIP Database")), side=1, line=3, adj=1, outer=FALSE)
```

**Figure 2. Trends in `r tmplist$county` County `r tmplist$fishery_name` effort by mode, for all seasons.** The dark blue line is the most likely number of trips and the light blue area shows the uncertainty. Absence of uncertainty can indicate sparse data.

### Summary
This summary shows that how the `r tmplist$fishery_name` effort in `r tmplist$county` County differs by mode over time and varies by season. This information should help inform the way fishing and especially certain modes of fishing are viewed, and it may also inform local governance decisions this `r tmplist$fishery_name`. 

